Global Optimization of the Analogue Method by Means of Genetic Algorithms

### 2016-01-21 Genetic algorithms - Journal of Climate/template.tex
Charle's corrections.

\abstract{The analogues method is based on a statistical relationship between synoptic atmospheric variables (predictors) and local weather (predictand), which we aim at forecasting. This relationship is expressed through many parameters that are usually calibrated by means of a semi-automatic sequential procedure. This calibration approach has strong limitations: it is made of successive steps and thus cannot handle parameters dependencies, and it cannot automatically optimize some parameters, such as the selection of the pressure levels and the temporal windows on which the predictors are compared.
In order to surpass these limitations, a global optimization technique was assessed, namely genetic algorithms, which can optimize jointly all parameters of the method and get closer to a global optimum by taking into account the parameters dependencies. Moreover, it can choose objectively parameters that were previously manually assessed, and can take into account new degrees of freedom that were unthinkable before.
These kind of optimization techniques need however to be tailored to the problem to solve. Multiple combinations of algorithms had to be assessed, and even new operators were developed, such as the \textit{chromosome of adaptive search radius} which was found to be very robust, in order to make recommendations for the use of Genetic Algorithms to optimize the analogue method. These recommendations are the main outcome of this work. It opens new perspective for the improvement of the analogue method, and its application to new regions or to new predictands.
}

@@

The analogue method (AM) relies on the hypothesis that similar situations in terms of atmospheric circulation are likely to lead to similar local weather \citep{Lorenz1956, Lorenz1969, Duband1970, Bontron2005}. The principle consists in sampling a certain number of past situations based on different atmospheric variables (predictors) in order to build a probabilistic forecast for a local weather variable of interest (predictand). The most common usage of the AM is for precipitation forecasting \citep[eg.][]{Guilbaud1997, Bontron2005, Bliefernicht2010, Marty2012, Horton2012, Radanovics2013, BenDaoud2015}, but AMs, or equivalent, were also used for short to medium term forecasting of daily temperatures \citep{Radinovic1975, Woodcock1980, Kruizinga1983}, wind \citep{Gordon1987}, snow avalanches \citep{Obled1980, Bolognesi1993}, insolation \citep{Bois1981}, and the trajectory of tropical cyclones \citep{Keenan1981, Sievers2000, Fraedrich2003}. Applications for monthly forecasts in many countries also exist, including Canada \citep{Shabbar1986},  Hungary \citep{Toth1989}, the Netherlands \citep{Nap1981}, and England \citep{Murray1974}, as well as seasonal forecasts: \citet{Barnett1978}, \citet{Bergen1982} and \citet{Livezey1988}.

The purpose of this paper is not to present in details the AM and its different parametrizations, which can be found in \citet{Horton2016} or \citet{BenDaoud2015}. The concepts one needs to understand here are the following: the AM consists in searching a certain number of past situations in a meteorological archive (typically reanalysis) that are the most similar, according to an analogy criteria, and to extract the observed values of the local weather variable of interest from another archive (typically local time series) in order to build the empirical distribution considered as the probabilistic forecast for the target day. Predictors can be multiple: for example the geopotential height at different pressure levels on different temporal windows (time of observation). The method is usually made of multiple levels of analogy, leading to successive subsampling on predictors of different nature (eg. atmospheric circulation, moisture variables, vertical motion, and air temperature).

Even though the method is rather simple compared to numerical weather prediction models, it contains a certain number of parameters one needs to determine, such as the choice of the predictor variable, its pressure level and temporal window to consider, the spatial domain to use for the comparison, as well as the analogy criteria itself, and finally the number of analogue situations to keep at each subsampling level.

The AM needs to be adapted to every new region it is applied, because the leading meteorological influences are specific per region. Even the selection of the pressure levels and the temporal windows should be reconsidered, when not the variable itself. For example, \citet{BenDaoud2010} found the vertical velocity relevant for the great plains in France, when \citet{Horton2012a} found no interest in this variable in an Alpine environment, because the vertical motion, which is mainly controlled by orographic effects, was already well related to the atmospheric circulation itself.

Up to now, the optimization of the method has been undertaken by semi-automatic sequential calibration procedures \citep[see][for its definition]{Bontron2004, Horton2016}. However, the choice of predictor variables, pressure levels and temporal windows were still to be chosen manually before optimizing the spatial windows and the number of analogues. Testing multiple combinations of these can be cumbersome, especially when considering multiple predictors within the same level of analogy. Thus, optimizing the method with the sequential technique is laborious, as many combinations of predictors (variables, pressure levels, temporal windows) have to be assessed. Moreover, the sequential technique proceeds to the optimization sequentially, ignoring potential dependencies between the parameters of the method, may they be within a single level of analogy or between them. However, any dependency between them cannot be excluded, which could lead us to select another parametrization if calibrated together. Simultaneous calibration of all parameters has never been undertaken so far. Thus, due to its sequential approach, the risk of ending in a local optimum is high and can not be avoided. Indeed, during the AM calibration it has been found that the resulting parameters may vary with initial choices (such as the number of analogues).

When creating the sequential calibration procedure, \citet{Bontron2004} was aware of the problem of dependencies between parameters and wrote: ''\textit{We perceive here the combinatorial aspect of our problem: variables and spatial windows are not independent. We will present our results by first searching the best variable [note: e.g. selection of the pressure level and the temporal window for the geopotential height] on a chosen spatial window, and next, the best window for the chosen variable. However, even by repeating the process, are we sure to obtain the optimal combination?}''. And later in his work: ''\textit{Our approach, which is again to vary the parameters one by one -- the others being fixed in a more or less arbitrary manner -- may therefore not exactly lead us to the optimal solution}''. \citet{Bliefernicht2010} has also faced the combinatorial issue of the parameters of the AM and concludes that one needs to be an expert to know their respective influence, their sensitivity and their nonlinear interactions. \citet{BenDaoud2010}, when calibrating the AM, also stated that ''\textit{the combinatory aspect related to the calibration was found to be too high for all the parameters to be calibrated simultaneously }''.

In order to overcome these limitations, two optimization techniques were assessed. First, \citet{Horton2012a} assessed the ability of the \citet{Nelder1965a} method based on a simplex approach. This technique did not provide satisfying results and failed at converging toward a unique solution. The parameter space of the AM is very complex and is not appropriate for a linear optimization technique. The conclusion was that global optimization techniques were necessary in order to calibrate AMs, as it is the only way to optimize all parameters of all analogy levels simultaneously. In addition, it can overcome the systematic manual assessments of all pressure levels and temporal windows. Finally, it can open new perspectives by allowing the addition of new degrees of freedom, such as a weighting of the criteria values between the pressure levels \citep[see][]{Horton2016b}, and the consideration of non-overlapping spatial windows between the pressure levels. The relevance of Genetic Algorithms (GAs) is presented here, which does not exclude that other global optimization techniques could also be successful.

This article is not about discussing the details of the results of an optimization with GAs, but describing how to use GAs in order to successfully optimize AMs. Indeed, GAs variants are numerous and they always need to be tailored to the problem to be addressed. This requires intensive and systematic comparisons of operators and options in order to identify the key factors leading the optimization and the respective sensitivity of the options. These analyses are presented here and result in recommendations for the use of GAs when applied to AMs. The demonstration of the benefit brought by such an approach in a specific case study is the topic of \citet{Horton2016b}.

We will begin by presenting the concepts of GAs as well as the assessed operators (section \ref{sec:gas}). The comparative analyses of the operators and their results are presented in section \ref{sec:assessment}, which lead us to the recommendations in section \ref{sec:recommendations}.

@@

GAs are part of the family of Evolutionary Algorithms \citep{Back1993b, Schwefel1993}, which get inspiration from some mechanisms of biological evolution, such as reproduction, genetic mutations, chromosomal crossovers, and natural selection. GAs are the most used technique from Evolutionary Algorithms \citep{Back1993b}, and they are constantly improving \citep{Haupt2004}. However, with time, the different methods of Evolutionary Algorithms tend to be similar and share many commonalities \citep{Back1996b, Haupt2004}.

@@

GAs mimic the evolution of a population of individuals in a new environment, by applying rules based on natural processes, such as DNA mutation, chromosomes crossover, natural selection, etc. Generation after generation, the DNA mixes and the strong genes cumulate in some individuals \citep{Beasley1996a}.

Applications of GAs are very diversified as they can handle many problems of various types \citep{Joines1996a}, even with very complex cost surfaces \citep{Haupt2004}. The objective function to optimize (often named fitness function in this context) can be of different types, only the resulting value is used for optimization. Indeed, these algorithms do not require any knowledge of the problem, but they must be adapted in order to perform optimally.

@@

A key element of the parametrization of GAs is finding the right balance between exploration and exploitation \citep{Back1992a, Smith1997a}. Exploration is characterized by a relatively high probability to assess the regions of the parameter space that have not yet been visited. This probability must be sufficiently large at the beginning of the optimization, so that the algorithm is capable of identifying the region where the global optimum is located. Exploitation is characterized by a local search in an area of interest, and generally makes small movements. The latter is interesting to refine the results at the end of the optimization.

@@

The GAs optimize a population of $N$ individuals (parameter sets). Each individual contains a chromosome (parameters of the AM in this case). Genes are the parameters constituting the chromosome. The floating-point representation \citep[instead of the canonical binary, see][]{Goldberg1989, Goldberg1990a, Holland1992b} of the genes was considered, as it was found more suited in multiple applications \citep{Michalewicz1996, Herrera1998a, Haupt2004, Back1996b, Gaffney2010a}.

There are numerous implementation variants of GAs, often optimal for a given problem \citep{Hart1991a, Schraudolph1992a}. The divergences are the operators implementation, through significantly different algorithms, which has an important effect on the results \citep{Gaffney2010a}. However, the structure of the method (Figure \ref{fig:structure_gas}) resulting from the work of \citet{Holland1992b} is common to most applications \citep{Back1993b}, and consists in the following steps:

\begin{enumerate}
	\item A population of $N$ individuals (parameter sets of the AM) is randomly generated, which constitutes the initial population.
	\item The fitness (performance score or objective function) of every individual is assessed.
	\item A natural selection is applied, after which only the best individuals remain, which constitutes the intermediate generation (IG).
	\item From this IG, couples are formed according to given rules.
	\item These couples proceed to reproduction, or chromosome crossover, in order to mix their genes (parameters) according to the operator version. New children are generated in order to fill the IG back to $N$ individuals.
	\item Parents and children are then subject to mutation, where some genes get affected, meaning some parameter values are randomly changed.
	\item The new formed generation is then re-assessed.
	\item According to the ending criteria, the optimization ends or starts again for another iteration.
\end{enumerate}

@@

The first step of the optimization is to generate the initial population. A population is a set of $N$ individuals (each of which represents a point in the space of potential solutions, a parameter set of the AM in this application) that are going to evolve. A generation is the population at a given time.

A random initialization based on a uniform distribution is the most current version. The size $N$ of the population is often a compromise between the computation time and the quality of the solution. $N$ must allow sufficient sampling of the solutions field \citep{Beasley1996a}, and should thus vary as a function of chromosome size (ie the number of parameters to be optimized).

@@

Natural selection is performed on the basis of the values of the objective function. The selection allows to only keep a certain part of the population, usually half ($N/2$), which can access the IG (with $N_{IG}$ members). If $N_{IG}$ is too high, the reproduction rate is too low, whereas if it is too small, the strong traits of individuals do not have the ability to accumulate in the same chromosome \citep{Haupt2004}. Several techniques exist, such as:

\begin{itemize}
	\item $N_{IG}$-elitism \citep{Michalewicz1996}: the population is sorted according to the value of the objective function and only the better half is preserved.

@@

\begin{itemize}
	\item Rank pairing: individuals are gathered in pairs according to their rank (classified on the performance scores). Consecutive ranks are put together (odd rows are associated with even rows). This approach is easy to achieve, but does not look like a natural process.

@@

	\item Tournament selection: This operator is similar to the one used in natural selection, but is applied here for the successive selection of each parent. To select a parent, a number of individuals (2 or 3) are randomly picked and the best is kept. This operation is performed twice, once for each partner. This approach imitates the breeding competition in nature \citep{Haupt2004}.
\end{itemize}

@@

Once the two parents are selected for breeding, they combine their chromosomes and produce two children, bringing the number of individuals in the population back to $N$ (the parents also return back in the total population in order to complement the next generation). The combination of chromosomes is carried out using a crossover operator, thereby generating two offspring having characteristics derived from both parents. It allows a mixing of genes and accumulation of positive mutations.

@@

\begin{itemize}
	\item Single-point crossover: a crossover point is randomly chosen for the pair. The genes (AM parameters) located after that point are exchanged in between the two chromosomes.

	\item Two-point crossover: works like the single-point crossover, but there are two intersections defining the segments to exchange. This approach, which significantly extends the search space for the children, is considered more efficient than the previous \citep{Beasley1993a}.

	\item Multiple-point crossover \citep{DeJong1975a}: it is a generalization of the previous, with a number of crossover points up to the number of genes.

@@

The evaluated and developed mutation operators are the following:

@@

	\item Non-uniform mutation \citep{Michalewicz1996}: two random numbers are picked based on a uniform law: $r_{1}$, which determines the direction of the change, and $r_{2}$, which determines its magnitude. The new value of the gene is given by the following equation:
	\begin{equation}
	g_{n}^{'} = 
	\left\lbrace \begin{array}{l l} 
	g_{n} + \left(b_{n}-g_{n}\right) r_{2} \left(1 - \dfrac{G}{G_{m}} \right)^{2} & if \; r_{1} < 0.5 \\
	g_{n} - \left(g_{n}-a_{n}\right) r_{2} \left(1 - \dfrac{G}{G_{m}} \right)^{2} & if \; r_{1} \geq 0.5 \\
	\end{array} \right.
	\label{equation_mutation_nonuniform_original}
	\end{equation}
	where $a_{n}$ is the is the lower bound of the $n$-th gene, $b_{n}$ its upper bound, $G$ the present generation, and $G_{m}$ the maximum number of generations.

@@

When the gene to mutate is represented by a list of distinct values (eg meteorological variable or analogy criterion), the random choice of a new value is always based on a uniform distribution, without notion of distance in the parameters space.

@@

A process of elitism was introduced on the natural selection as well as on mutations. This ensures the survival of the best individual so that a better solution is never lost. After the natural selection operator, if the best individual has not been selected, it is copied to the IG instead of an individual randomly picked. After mutation, if the best individual has mutated and if its new version has a lower performance score than the original, the latter is also reinserted in the IG instead of an individual randomly chosen.

@@

Some constraints need to be taken into account. For example, when a crossover or a mutation operation results in a parameter value standing out of the authorized bounds, it has to be brought back within the limits. Moreover, the parameters are of different nature: some are continuous, some are discretized, and finally, some are independent elements in an array, such as the selection of the meteorological variable. New values resulting from the optimizer need to respect the type of data it represents.

@@

The GAs parametrization, such as the mutation rate, population size, natural selection options, and so on, is difficult given the high number of existing variants, each developed for a specific problem \citep{Haupt2004, Costa2007a}. The parametrization depends on the objective function, implementation variants, the range of the parameters to be optimized, and performance indicators. Thus, different studies suggest very different parametrization.

\citet{DeJong1975a} and \citet{Grefenstette1986} compared different implementations and parametrizations of GAs on functions of varying complexity. They observed that a small population size improves the initial performance, while a large population improves long-term performance. They also observed that the ratio of the population to keep for the IG is around 50\% (45\% to 60\%).

Values of the mutation rate varies broadly between the studies: from 0.001 \citep{DeJong1975a} to 0.2 \citep{Haupt2004}. \citet{Back1996b} showed that mutation rates higher than the usual ranges are more optimal at the beginning of the optimization, allowing further exploration. Incremental approaches with varying mutation rates are certainly more optimal but more complex to implement \citep{Back1996a, Back1996b}.

@@

One of the goals being to make recommendations of parametrization in order to optimize the AM, a systematic procedure was adopted. The results are summarized hereafter \citep[see][for the details]{Horton2012a}. Concepts were taken from the factorial design approach \citep[see eg.][]{Costa2005a,Costa2007a,Mariano2010a}, which is sometimes used for comparative analysis of different parametrizations of GAs. It allows isolating the effect of a parameter under different combinations of the other options. A procedure by stages was adopted, analyzing in details and in a systematic way every variants of the implemented operators, in combination with multiple other options and parameters in order to take into account eventual co-dependencies. The goal here is not to focus on the performance score obtained through optimization, neither the values of the new optimized parameters \citep[covered in][]{Horton2016b}, but to explain how to use GAs to optimize AMs in an efficient way.

In order to evaluate a combination of operators/options, 10 optimizations per parametrization of GAs were processed. The performances were characterized by four indicators: (i) mean performance score: average of the final scores of the 10 optimizations, (ii) convergence: the number of optimizations that converged to a supposed global optimum, (iii) number of generations: characterization of the convergence speed, and (iv) number of evaluations of the objective function: characterization of the required calculation time.

@@

After a first overview of the results, GAs have quicky been proven successful at optimizing the AM. Indeed, the performance score of the sequential approach has been quickly exceeded without adding any new parameter to the method (Figure \ref{fig:gas_evolution_good}). Even GAs parametrization that will be further considered as inadequate (Figure \ref{fig:gas_evolution_bad}) did significantly better that the sequential approach. The amplitude of the improvement is not the main outcome here. The most important point is that GAs are proven successful at optimizing AMs automatically, globally, and objectively.

@@

The results illustrate the effect of an operator when its contribution is isolated from the other operators. It means that we compare the effect of the studied operator when all the other operators are equivalent, while assessing multiple combinations. This contribution is then summarized relatively to the other options, as a percentage of gain/loss regarding the mean of all variants. For example, to evaluate the performance of the uniform crossover operator, its performance is compared to the average of all crossover operators while retaining the same population size, the same mutation operators, natural selection, and selection of couples.

@@

Analysis of crossover operators (Figure \ref{fig:operator_crossover_score}) reveal some slightly superior options, some inappropriate, and many averages. Binary-like crossover (especially with 2 points of intersection, whether $\beta$ is shared or not) are significantly better than the others, especially in terms of convergence (not shown). Others operators can also be considered usable.

>>>>

\abstract{The Analogue Method is based on a statistical relationship between synoptic atmospheric variables (predictors) and local weather (predictand), which we aim at forecasting. This relationship is expressed through many parameters that are usually calibrated by means of a semi-automatic sequential procedure. This calibration approach has strong limitations: it is made of successive steps and thus cannot handle parameters dependencies, and it cannot automatically optimize some parameters, such as the selection of the pressure levels and the temporal windows on which the predictors are compared.
In order to surpass these limitations, a global optimization technique was assessed, namely genetic algorithms, which can optimize jointly all parameters of the method and get closer to a global optimum by taking into account the parameters dependencies. Moreover, it can choose objectively parameters that were previously manually assessed, and can take into account new degrees of freedom that were unthinkable before.
These kind of optimization techniques need however to be tailored to the problem to solve. Multiple combinations of algorithms had to be assessed, and even new operators were developed, such as the \textit{chromosome of adaptive search radius} which was found to be very robust, in order to make recommendations for the use of Genetic Algorithms to optimize the analogue method. These recommendations are the main outcome of this work. It opens new perspective for the improvement of the analogue method, and its application to new regions or to new predictands.
}

@@

The Analogue Method (AM) relies on the hypothesis that similar situations in terms of atmospheric circulation are likely to lead to similar local weather \citep{Lorenz1956, Lorenz1969, Duband1970, Bontron2005}. The principle consists in sampling a certain number of past situations based on different atmospheric variables (predictors) in order to build a probabilistic forecast for a local weather variable of interest (predictand). The most common usage of the AM is for precipitation forecasting \citep[eg.][]{Guilbaud1997, Bontron2005, Bliefernicht2010, Marty2012, Horton2012, Radanovics2013, BenDaoud2015}, but AMs, or equivalent, were also used for short to medium term forecasting of daily temperatures \citep{Radinovic1975, Woodcock1980, Kruizinga1983}, wind \citep{Gordon1987}, snow avalanches \citep{Obled1980, Bolognesi1993}, insolation \citep{Bois1981}, and the trajectory of tropical cyclones \citep{Keenan1981, Sievers2000, Fraedrich2003}. Applications for monthly forecasts in many countries also exist, including Canada \citep{Shabbar1986},  Hungary \citep{Toth1989}, the Netherlands \citep{Nap1981}, and England \citep{Murray1974}, as well as seasonal forecasts: \citet{Barnett1978}, \citet{Bergen1982} and \citet{Livezey1988}.

The purpose of this paper is not to present in details the AM and its different parameterizations, which can be found in \citet{Horton2016} or \citet{BenDaoud2015}. The concepts one needs to understand here are the following: the AM consists in searching a certain number of past situations in a meteorological archive (typically reanalysis datasets) that are the most similar, according to an analogy criteria, and to extract the observed values of the local weather variable of interest from another archive (typically local or basin averaged time series) in order to build the conditional empirical distribution considered as the probabilistic forecast for the target day. Predictors can be varied: for example the geopotential height at different pressure levels on different temporal windows (time of observation). The method is usually made of several levels of analogy, leading to successive subsampling on predictors of different nature (eg. atmospheric circulation, moisture variables, vertical motion, and air temperature).

Even though the method is rather simple compared to numerical weather prediction (NWP) models, it contains a certain number of parameters one needs to determine, such as the choice of the predictor variable, its pressure level and temporal window to consider, the spatial domain to use for the comparison, as well as the analogy criteria itself, and finally the number of analogue situations to keep at each subsampling level.

The AM needs to be adapted to every new region considered, because the leading meteorological influences may be specific per region. Even the selection of the pressure levels and the temporal windows should be reconsidered, when not the predictor variable itself. For example, \citet{BenDaoud2010} found the vertical velocity relevant for the great plains in France, when \citet{Horton2012a} found it useless in an Alpine environment, because the vertical motion, which is mainly controlled by orographic effects, was already well related to the atmospheric circulation itself.

Up to now, the optimization of the method has been undertaken by semi-automatic sequential calibration procedures \citep[see][for its definition]{Bontron2004, Horton2016}. However, the selection of predictor variables, pressure levels and temporal windows had still to be made manually before optimizing the spatial windows and the number of analogues. Testing multiple combinations of these is very combinatorial and becomes quickly cumbersome, especially when considering multiple predictors within the same level of analogy. Thus, optimizing the method with the sequential technique is laborious, as many combinations of predictors (variables, pressure levels, temporal windows) have to be assessed. Moreover, proceeding to the optimization sequentially ignores potential dependencies between the parameters of the method, may they be within a single level of analogy or between them. However, such dependencies between them cannot be excluded, which could lead to another parametrization if calibrated together. However, simultaneous calibration of all parameters has never been undertaken so far. Thus, due to the sequential approach, the risk of ending in a local optimum is high and can not be avoided. Indeed, during the AM calibration it has been found that the resulting parameters may vary with initial choices (such as the number of analogues).

When creating the sequential calibration procedure, \citet{Bontron2004} was aware of the problem of dependencies between parameters and wrote: ''\textit{We perceive here the combinatorial aspect of our problem: variables and spatial windows are not independent. We will present our results by first searching the best variable [note: e.g. selection of the pressure level and the temporal window for the geopotential height] on a chosen spatial window, and next, the best window for the chosen variable. However, even by repeating the process, are we sure to obtain the optimal combination?}''. And later in his work: ''\textit{Our approach, which is again to vary the parameters one by one -- the others being fixed in a more or less arbitrary manner -- may therefore not exactly lead us to the optimal solution}''. \citet{Bliefernicht2010} has also faced the combinatorial issue of the parameters of the AM and concludes that one needs to be an expert to have a sense of their respective influence, sensitivity and nonlinear interactions. \citet{BenDaoud2010}, when calibrating the AM, also stated that ''\textit{the combinatory aspect related to the calibration was found to be too high for all the parameters to be calibrated simultaneously }''.

In order to overcome these limitations, two optimization techniques were assessed. First, \citet{Horton2012a} assessed the ability of the \citet{Nelder1965a} method based on a simplex approach. This technique did not provide satisfying results and failed at converging toward a unique solution. The parameter space of the AM is very complex and is inappropriate for a linear optimization technique. The conclusion was that global optimization techniques were necessary in order to calibrate AMs, as it is the only way to optimize all parameters of all analogy levels simultaneously. In addition, it can overcome the systematic manual assessments of all pressure levels and temporal windows. Finally, it can open new perspectives by allowing the addition of new degrees of freedom, such as a weighting of the criteria values between the pressure levels \citep[see][]{Horton2016b}, and the consideration of differentiated spatial windows between the pressure levels. The relevance of Genetic Algorithms (GAs) is presented here, which does not exclude that other global optimization techniques could eventually be successful.

This article is not about discussing the details of the results of an optimization with GAs, but describing how GAs are to be used in order to successfully optimize AMs. Indeed, GAs variants are numerous and always need to be tailored to the problem addressed. This requires intensive and systematic comparisons of operators and options in order to identify the key factors leading the optimization and the respective sensitivity of the options. Such analyses are presented here and will result in recommendations for the use of GAs when applied to AMs. The demonstration of the benefit brought by such an approach on a specific case study is the topic of \citet{Horton2016b}.

We will begin by presenting the concepts of GAs as well as the assessed operators (operations performed on the AM parameters; see section \ref{sec:gas}). The comparative analyses of the operators and their results are presented in section \ref{sec:assessment}, which lead us to the recommendations in section \ref{sec:recommendations}.

@@

GAs are part of the family of Evolutionary Algorithms \citep{Back1993b, Schwefel1993}, inspired by some mechanisms of biological evolution, such as reproduction, genetic mutations, chromosomal crossovers, and natural selection. GAs are the most used technique among Evolutionary Algorithms \citep{Back1993b}, and they are constantly improving \citep{Haupt2004}. However, with time, the different methods of Evolutionary Algorithms tend to be similar and share many commonalities \citep{Back1996b, Haupt2004}.

@@

GAs mimic the evolution of a population of individuals in a new environment, by applying rules based on natural processes, such as DNA mutation, chromosomes crossover, natural selection, etc. Generation after generation, the DNA mixes and the best suited, or best adapted, genes cumulate in some individuals \citep{Beasley1996a}.

Applications of GAs are diversified and can handle problems of various types \citep{Joines1996a}, even with very complex cost surfaces \citep{Haupt2004}. The objective function to optimize (often named fitness function in this context) can be of different types, but GAs must be adapted in order to perform optimally.

@@

A key element of the parametrization of GAs is finding the right balance between exploration and exploitation \citep{Back1992a, Smith1997a}. Exploration is characterized by a relatively high probability to assess the regions of the parameter space that have not yet been visited. This probability must be sufficiently large at the beginning of the optimization, so that the algorithm is capable of identifying the region where the global optimum is likely located. Exploitation is characterized by a local search in an area of interest, and generally makes small movements. The latter is interesting to refine the results at the end of the optimization.

@@

The GAs optimize a population of $N$ individuals (parameter sets). Each individual contains a chromosome (set of parameters of the AM in this case). Genes are the individual parameters constituting the chromosome. They can be either categorical (e.g. geopotential, or temperature), discrete (e.g. number of analogues to select, from 1 to 40), or continuous. The floating-point representation \citep[instead of the canonical binary, see][]{Goldberg1989, Goldberg1990a, Holland1992b} of the genes was considered, as it was found more suited in multiple applications \citep{Michalewicz1996, Herrera1998a, Haupt2004, Back1996b, Gaffney2010a}.

There are numerous implementation variants of GAs, often optimal for a given problem \citep{Hart1991a, Schraudolph1992a}. The divergences are in the operators implementation, through significantly different algorithms, which has an important effect on the results \citep{Gaffney2010a}. Here, operators are defined as the mechanisms that modify the values of the genes to try bringing individuals (or chromosomes) closer to an optimum of the fitness function. The structure of the method (Figure \ref{fig:structure_gas}) resulting from the work of \citet{Holland1992b} is common to most applications \citep{Back1993b}, and consists in the following steps:

\begin{enumerate}
	\item A population of $N$ individuals (parameter sets of the AM) is randomly generated, which constitutes the initial population.
	\item The fitness (performance score or objective function) of every individual is assessed.
	\item A natural selection is applied, after which only the best individuals remain, which constitutes the intermediate generation (IG).
	\item From this IG, couples are formed according to given rules.
	\item These couples proceed to reproduction, or chromosome crossover, in order to mix their genes (parameters) according to the selected operator version. New children are generated in order to refill the IG back to $N$ individuals.
	\item Parents and children are then subject to mutation, where some genes get affected, meaning some parameter values are randomly changed.
	\item The new formed generation is then re-assessed.
	\item According to the ending criteria, the optimization ends or starts again for another iteration.
\end{enumerate}

@@

The first step of the optimization is to generate an initial population. A population is a set of $N$ individuals (each of which represents, in this application, a parameter set of the AM and thus a point in the space of potential solutions) that are going to evolve. A generation is the population considered at a given time.

A random initialization based on a uniform sampling is the most current version. The size $N$ of the population is often a compromise between the computation time and the quality of the solution. $N$ must allow sufficient sampling of the solutions field \citep{Beasley1996a}, and should thus vary as a function of chromosome size (ie the number of genes or parameters to be optimized).

@@

Natural selection is performed on the basis of the values of the objective function. The selection allows to only keep a certain part of the population, usually half ($N/2$), which can access the IG (with $N_{IG}$ members). If $N_{IG}$ is too high, the reproduction rate is too low, whereas if it is too small, the strong traits of individuals will not have the ability to accumulate in the same chromosome \citep{Haupt2004}. Several techniques exist, such as:

\begin{itemize}
	\item $N_{IG}$-elitism \citep{Michalewicz1996}: the population is ranked according to the value of the objective function and only the better half is preserved.

@@

\begin{itemize}
	\item Rank pairing: individuals are gathered in pairs according to their rank (based on the performance scores). Consecutive ranks are put together (odd rows are associated with even rows). This approach is easy to achieve, but does not look like a natural process.

@@

	\item Tournament selection: This operator is similar to the one used in natural selection, but is applied here for the successive selection of each parent. To select a parent, a number of individuals (2 or 3) are randomly picked and the best is kept. This operation is performed twice, once for each partner and imitates the breeding competition in nature \citep{Haupt2004}.
\end{itemize}

@@

Once the two parents are selected for breeding, they combine their chromosomes and produce two children, bringing the number of individuals in the population back to $N$ (the parents also return back in the total population in order to complement the next generation). The combination of chromosomes is carried out using a crossover operator, thereby generating two offspring having characteristics derived from both parents. It allows a mixing of genes and a potential accumulation of positive mutations.

@@

\begin{itemize}
	\item Single-point crossover: within the chromosome, a crossover point is randomly chosen for the pair. The genes (AM parameters) located after that point are exchanged in between the two chromosomes.

	\item Two-point crossover: works like the single-point crossover, but there are two intersections defining the segments to be exchanged. This approach, which significantly extends the search space for the children, is considered more efficient than the previous \citep{Beasley1993a}.

	\item Multiple-point crossover \citep{DeJong1975a}: a generalization of the previous, with a number of crossover points up to the number of genes.

@@

The evaluated and developed mutation operators are listed hereafter. They apply to genes made of continuous or discrete variables, but not categorical (eg meteorological variable or analogy criterion). In the latter case, the random choice of a new value is always based on a uniform distribution, without notion of distance in the parameters space.

@@

	\item Non-uniform mutation \citep{Michalewicz1996}: two random numbers are picked based on a uniform law: $r_{1}$, which determines the direction of the change, and $r_{2}$, which determines its magnitude. The new value of the gene is given by the following equation, according to a predefined number of generations:
	\begin{equation}
	g_{n}^{'} = 
	\left\lbrace \begin{array}{l l} 
	g_{n} + \left(b_{n}-g_{n}\right) r_{2} \left(1 - \dfrac{G}{G_{m}} \right)^{2} & if \; r_{1} < 0.5 \\
	g_{n} - \left(g_{n}-a_{n}\right) r_{2} \left(1 - \dfrac{G}{G_{m}} \right)^{2} & if \; r_{1} \geq 0.5 \\
	\end{array} \right.
	\label{equation_mutation_nonuniform_original}
	\end{equation}
	where $a_{n}$ is the is the lower bound of the $n$-th gene, $b_{n}$ its upper bound, $G$ the present generation, and $G_{m}$ the maximum number of generations.

@@

@@

A process of elitism was introduced on the natural selection as well as on mutations. This ensures the survival of the best individual so that a better solution is never lost. After the natural selection operator, if the previously best individual has not been selected, it is copied to the IG instead of an individual randomly picked. After mutation, if the previously best individual has mutated and if its new version has a lower performance score than the original, the latter is also reinserted in the IG instead of an individual randomly chosen.

@@

Some constraints need to be taken into account. For example, when a crossover or a mutation operation results in a parameter value standing out of the authorized bounds, it has to be brought back within the limits. Moreover, the parameters are of different nature: some are continuous, some are discrete, and finally, some are categorical, i.e. independent elements in an array, such as the selection of the meteorological variable. New values resulting from the optimizer need to respect the type of data it represents.

@@

The GAs parametrization, i.e. the a priori choices such as the mutation rate, population size, natural selection options, etc, appears difficult given the high number of existing variants, each developed for a specific problem \citep{Haupt2004, Costa2007a}. The parametrization depends on the objective function, implementation variants, the range of the parameters to be optimized, and performance indicators. Thus, different studies suggest very different parameterizations.

\citet{DeJong1975a} and \citet{Grefenstette1986} compared different implementations and parametrizations of GAs on functions of varying complexity. They observed that a small population size improves the initial performance, while a large population improves latter performance. They also observed that the proportion of the population to keep for the IG is around 50\% (45\% to 60\%).

Values of the mutation rate varies broadly between studies: from 0.001 \citep{DeJong1975a} to 0.2 \citep{Haupt2004}. \citet{Back1996b} showed that mutation rates higher than the usual ranges are more optimal at the beginning of the optimization, allowing further exploration. Varying mutation rates are certainly more optimal but more complex to implement \citep{Back1996a, Back1996b}.

@@

One of our goals being to make recommendations of parametrization in view of optimizing the AM, a systematic procedure was adopted. The results are summarized hereafter \citep[see][for the details]{Horton2012a}. Concepts were taken from the factorial design approach \citep[see eg.][]{Costa2005a,Costa2007a,Mariano2010a}, which is sometimes used for comparative analysis of different parametrizations of GAs. It allows isolating the effect of a parameter under different combinations of the other options. A procedure by stages was adopted, analyzing in details and in a systematic way every variants of the implemented operators, in combination with multiple other options and parameters in order to take into account eventual co-dependencies. The goal here is not to focus on the performance score obtained through optimization, neither the values of the new optimized parameters \citep[covered in][]{Horton2016b}, but to explain how to use GAs to optimize AMs in an efficient way.

In order to evaluate a combination of operators/options, 10 optimizations per parametrization of GAs were processed. Such assessment is not possible on the whole archive length, and had to be performed on a reduced period. The performances were characterized by four indicators: (i) mean performance score: average of the final scores of the 10 optimizations, (ii) convergence: the number of optimizations that converged to a supposed global optimum, (iii) number of generations: characterization of the convergence speed, and (iv) number of evaluations of the objective function: characterization of the required calculation time.

@@

After a first overview of the results, GAs have quicky proved successful at optimizing the AM. The performance score of the sequential approach has been quickly exceeded without adding any new parameter to the method (Figure \ref{fig:gas_evolution_good}). Even GAs parameterizations that will be considered later as inadequate (Figure \ref{fig:gas_evolution_bad}) did significantly better that the sequential approach. The amplitude of the improvement is not the main outcome here. The most important point is that GAs proved successful at optimizing AMs automatically, globally, and objectively.

@@

The results illustrate the effect of an operator when its contribution is isolated from the other operators. It means that we analyze the effect of a given operator for equivalent conditions (same settings of other operators). Multiple combinations with other operators are assessed. This contribution is then summarized as a percentage of gain/loss regarding the mean of all variants, for equivalent external conditions. For example, to evaluate the performance of the uniform crossover operator, its performance is compared to the average of all crossover operators while retaining the same population size, the same mutation operators, natural selection, and selection of couples.

@@

Analysis of crossover operators (Figure \ref{fig:operator_crossover_score}) reveal some slightly superior options, some inappropriate, and many average. Binary-like crossover (especially with 2 points of intersection, whether $\beta$ is shared or not) are significantly better than the others, especially in terms of convergence (not shown). Others operators can also be considered usable.
